---
title: "2017 New Zealand Election"
output: 
  flexdashboard::flex_dashboard:
    theme: spacelab
    self_contained: false
    lib_dir: lib
    mathjax: null
    navbar:
      - { title: "Intro", icon: "fa-home", href: "index.html", align: left }
      - { title: "Dumbbells", icon: "ion-ios-settings-strong", href: "dumbbells.html", align: left }
      - { title: "Hexmaps", icon: "ion-cube", href: "hexmaps.html", align: left }
      - { title: "3D Hexmaps", icon: "fa-cubes", href: "", align: left }
      - { title: "@dakvid", icon: "fa-twitter", href: "https://twitter.com/dakvid", align: right }
    social: menu
    source_code: https://github.com/dakvid/election2017
---

```{r init, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```

```{r setup, include=FALSE}
library(flexdashboard)
library(sf)
library(magrittr)
library(dplyr)
library(purrr)
library(stringr)
library(plotly)

load("data_geo/nzhex.rda")
load("data_votes/GE2017.rda")

party_votes <-
  GE2017 %>% 
  filter(VotingType == "Party") %>% 
  group_by(id, Party) %>% 
  summarise(Votes = sum(Votes)) %>% 
  ungroup() %>% 
  filter(!str_detect(Party, "nformal"))
```


```{r 3d_fn}
draw_3dhex <- function(mesh, largs) {
  hex_x <- st_coordinates(largs$geometry)[, "X"]
  hex_y <- st_coordinates(largs$geometry)[, "Y"]
  mesh %>% 
    add_mesh(
      x = ~rep(-hex_x, 2),
      y = ~rep(-hex_y, 2),
      z = ~c(rep(0, 7), rep(largs$Votes, 7)),
      name = ~paste(largs$electorate_name, prettyNum(largs$Votes, big.mark = ','), "votes"),
      hoverinfo = "name"
    )
}

xyaxis <- 
  list(
    title = "",
    showgrid = FALSE,
    showticklabels = FALSE,
    ticks = ""
  )
```


## Explanation {.sidebar}

Click and drag to rotate the plot in three dimensions!

The columns are the same hexagons as in the
[flat hexmaps](hexmaps.html)
but the heights are the numbers of votes received.

Before you ask: the colours are randomly assigned (by default).
I had a bit of trouble trying to get these working,
and still haven't figured out how to unsquash the hexagons or
assign a uniform colour.

These views allow us to see a few things more clearly,
such as how much higher the Green's support is in Wellington,
and the variation in support for the Māori Party among the
Māori electorates.



## 3D Hexmaps {.tabset .tabset-fade}

### National

```{r 3dhex_national}
party_votes %>% 
  filter(Party == "National Party") %>% 
  inner_join(nzhex) %>% 
  select(electorate_name, Votes, geometry) %>% 
  transpose() %>% 
  reduce(draw_3dhex, .init = plot_ly(type = "mesh3d", alphahull = 0)) %>% 
  layout(
    scene = list(
      xaxis = xyaxis,
      yaxis = xyaxis,
      zaxis = list(title = "Votes")
    ),
    aspectratio = c(5, 1, 1),
    title = "2017 Votes for the National Party"
  )
```

### Labour

```{r 3dhex_labour}
party_votes %>% 
  filter(Party == "Labour Party") %>% 
  inner_join(nzhex) %>% 
  select(electorate_name, Votes, geometry) %>% 
  transpose() %>% 
  reduce(draw_3dhex, .init = plot_ly(type = "mesh3d", alphahull = 0)) %>% 
  layout(
    scene = list(
      xaxis = xyaxis,
      yaxis = xyaxis,
      zaxis = list(title = "Votes")
    ),
    title = "2017 Votes for the Labour Party"
  )
```


### NZ First

```{r 3dhex_nzfirst}
party_votes %>% 
  filter(Party == "New Zealand First Party") %>% 
  inner_join(nzhex) %>% 
  select(electorate_name, Votes, geometry) %>% 
  transpose() %>% 
  reduce(draw_3dhex, .init = plot_ly(type = "mesh3d", alphahull = 0)) %>% 
  layout(
    scene = list(
      xaxis = xyaxis,
      yaxis = xyaxis,
      zaxis = list(title = "Votes")
    ),
    title = "2017 Votes for the NZ First Party"
  )
```


### Green

```{r 3dhex_green}
party_votes %>% 
  filter(Party == "Green Party") %>% 
  inner_join(nzhex) %>% 
  select(electorate_name, Votes, geometry) %>% 
  transpose() %>% 
  reduce(draw_3dhex, .init = plot_ly(type = "mesh3d", alphahull = 0)) %>% 
  layout(
    scene = list(
      xaxis = xyaxis,
      yaxis = xyaxis,
      zaxis = list(title = "Votes")
    ),
    title = "2017 Votes for the Green Party"
  )
```


### Māori

```{r 3dhex_maori}
party_votes %>% 
  filter(Party == "Māori Party") %>% 
  inner_join(nzhex) %>% 
  select(electorate_name, Votes, geometry) %>% 
  transpose() %>% 
  reduce(draw_3dhex, .init = plot_ly(type = "mesh3d", alphahull = 0)) %>% 
  layout(
    scene = list(
      xaxis = xyaxis,
      yaxis = xyaxis,
      zaxis = list(title = "Votes")
    ),
    title = "2017 Votes for the Māori Party"
  )
```


### TOP

```{r 3dhex_top}
party_votes %>% 
  filter(Party == "The Opportunities Party (TOP)") %>% 
  inner_join(nzhex) %>% 
  select(electorate_name, Votes, geometry) %>% 
  transpose() %>% 
  reduce(draw_3dhex, .init = plot_ly(type = "mesh3d", alphahull = 0)) %>% 
  layout(
    scene = list(
      xaxis = xyaxis,
      yaxis = xyaxis,
      zaxis = list(title = "Votes")
    ),
    title = "2017 Votes for The Opportunities Party"
  )
```



### ACT

```{r 3dhex_act}
party_votes %>% 
  filter(Party == "ACT New Zealand") %>% 
  inner_join(nzhex) %>% 
  select(electorate_name, Votes, geometry) %>% 
  transpose() %>% 
  reduce(draw_3dhex, .init = plot_ly(type = "mesh3d", alphahull = 0)) %>% 
  layout(
    scene = list(
      xaxis = xyaxis,
      yaxis = xyaxis,
      zaxis = list(title = "Votes")
    ),
    title = "2017 Votes for the ACT Party"
  )
```





<script>
  (function(i,s,o,g,r,a,m){i['GoogleAnalyticsObject']=r;i[r]=i[r]||function(){
  (i[r].q=i[r].q||[]).push(arguments)},i[r].l=1*new Date();a=s.createElement(o),
  m=s.getElementsByTagName(o)[0];a.async=1;a.src=g;m.parentNode.insertBefore(a,m)
  })(window,document,'script','https://www.google-analytics.com/analytics.js','ga');

  ga('create', 'UA-30013353-2', 'auto');
  ga('send', 'pageview');

</script>
