---
title: "2017 New Zealand Election"
output: 
  flexdashboard::flex_dashboard:
    theme: spacelab
    self_contained: false
    lib_dir: lib
    mathjax: null
    navbar:
      - { title: "Intro", icon: "fa-home", href: "index.html", align: left }
      - { title: "Dumbbells", icon: "ion-ios-settings-strong", href: "", align: left }
      - { title: "Hexmaps", icon: "ion-cube", href: "hexmaps.html", align: left }
      - { title: "3D Hexmaps", icon: "fa-cubes", href: "3dhex.html", align: left }
      - { title: "@dakvid", icon: "fa-twitter", href: "https://twitter.com/dakvid", align: right }
    social: menu
    source_code: https://github.com/dakvid/election2017
---

```{r init, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```

```{r setup, include=FALSE}
library(flexdashboard)
library(sf)
library(magrittr)
library(dplyr)
library(forcats)
library(purrr)
library(stringr)
library(leaflet)
library(leaflet.minicharts)
library(nzelect)
library(htmltools)
library(htmlwidgets)
library(plotly)

load("data_geo/nzhex.rda")
load("data_votes/GE2017.rda")
data("GE2014")
GE2014 %<>%
  mutate(id = Electorate %>% str_extract("[0-9]+") %>% as.integer(),
         Party = if_else(Party == "Maori Party", "Māori Party", Party))

party_voters_2017 <-
  GE2017 %>% 
  filter(VotingType == "Party",
         !str_detect(Party, "nformal")) %>% 
  group_by(id) %>% 
  summarise(Voters = sum(Votes)) %>% 
  ungroup()

candidate_voters_2017 <-
  GE2017 %>% 
  filter(VotingType == "Candidate",
         !str_detect(Candidate, "nformal")) %>% 
  group_by(id) %>% 
  summarise(Voters = sum(Votes)) %>% 
  ungroup()

party_voters_2014 <-
  GE2014 %>% 
  filter(VotingType == "Party",
         !str_detect(Party, "nformal")) %>% 
  group_by(id) %>% 
  summarise(Voters2014 = sum(Votes)) %>% 
  ungroup()

candidate_voters_2014 <-
  GE2014 %>% 
  filter(VotingType == "Candidate",
         !str_detect(Candidate, "nformal")) %>% 
  group_by(id) %>% 
  summarise(Voters2014 = sum(Votes)) %>% 
  ungroup()
```


# 

```{r db_data}
party_votes <-
  GE2017 %>% 
  filter(VotingType == "Party") %>% 
  group_by(id, Party) %>% 
  summarise(Votes = sum(Votes)) %>% 
  ungroup() %>% 
  inner_join(party_voters_2017) %>% 
  mutate(PerCent = round(Votes / Voters * 100, 1))

party_votes_2014 <-
  GE2014 %>% 
  filter(VotingType == "Party") %>% 
  group_by(id, Party) %>% 
  summarise(Votes2014 = sum(Votes)) %>% 
  ungroup() %>% 
  inner_join(party_voters_2014) %>% 
  mutate(PerCent2014 = round(Votes2014 / Voters2014 * 100, 1))

pv_change <-
  inner_join(
    party_votes %>% 
      inner_join(nzhex) %>% 
      select(Party, Electorate = electorate_name, `2017` = PerCent),
    party_votes_2014 %>% 
      inner_join(nzhex) %>% 
      select(Party, Electorate = electorate_name, `2014` = PerCent2014)
  )
```


## Plots {.tabset}

### National

```{r db_national}
pv_change %>% 
  filter(Party == "National Party") %>% 
  mutate(Electorate = fct_reorder(Electorate, `2017`)) %>% 
  plot_ly(color = I("gray80")) %>% 
  add_segments(
    x = ~pmin(`2014`, `2017`), xend = ~pmax(`2014`, `2017`),
    y = ~Electorate, yend = ~Electorate,
    showlegend = FALSE
  ) %>% 
  add_markers(
    x = ~`2014`, y = ~Electorate, name = "2014",
    color = I("grey"), hoverinfo = "text",
    text = ~paste0(`2014`, "% of ", Electorate, " votes in 2014")
  ) %>% 
  add_markers(
    x = ~`2017`, y = ~Electorate, name = "2017",
    color = I(parties_v["National"]), hoverinfo = "text",
    text = ~paste0(`2017`, "% of ", Electorate, " votes in 2017")
  ) %>% 
  layout(
    title = "Relative Votes for the National Party by Electorate",
    xaxis = list(title = "Votes (Percent)"),
    yaxis = list(title = "", dtick = 1),
    margin = list(l = 130),
    legend = list(x = 0.8, y = 0.1)
  )
```

### Labour

```{r db_labour}
pv_change %>% 
  filter(Party == "Labour Party") %>% 
  mutate(Electorate = fct_reorder(Electorate, `2017`)) %>% 
  plot_ly(color = I("gray80")) %>% 
  add_segments(
    x = ~pmin(`2014`, `2017`), xend = ~pmax(`2014`, `2017`),
    y = ~Electorate, yend = ~Electorate,
    showlegend = FALSE
  ) %>% 
  add_markers(
    x = ~`2014`, y = ~Electorate, name = "2014",
    color = I("grey"), hoverinfo = "text",
    text = ~paste0(`2014`, "% of ", Electorate, " votes in 2014")
  ) %>% 
  add_markers(
    x = ~`2017`, y = ~Electorate, name = "2017",
    color = I(parties_v["Labour"]), hoverinfo = "text",
    text = ~paste0(`2017`, "% of ", Electorate, " votes in 2017")
  ) %>% 
  layout(
    title = "Relative Votes for the Labour Party by Electorate",
    xaxis = list(title = "Votes (Percent)"),
    yaxis = list(title = "", dtick = 1),
    margin = list(l = 130),
    legend = list(x = 0.8, y = 0.1)
  )
```

### NZ First

```{r db_nzf}
pv_change %>% 
  filter(Party == "New Zealand First Party") %>% 
  mutate(Electorate = fct_reorder(Electorate, `2017`)) %>% 
  plot_ly(color = I("gray80")) %>% 
  add_segments(
    x = ~pmin(`2014`, `2017`), xend = ~pmax(`2014`, `2017`),
    y = ~Electorate, yend = ~Electorate,
    showlegend = FALSE
  ) %>% 
  add_markers(
    x = ~`2014`, y = ~Electorate, name = "2014",
    color = I("grey"), hoverinfo = "text",
    text = ~paste0(`2014`, "% of ", Electorate, " votes in 2014")
  ) %>% 
  add_markers(
    x = ~`2017`, y = ~Electorate, name = "2017",
    color = I(parties_v["NZ First"]), hoverinfo = "text",
    text = ~paste0(`2017`, "% of ", Electorate, " votes in 2017")
  ) %>% 
  layout(
    title = "Relative Votes for the NZ First Party by Electorate",
    xaxis = list(title = "Votes (Percent)"),
    yaxis = list(title = "", dtick = 1),
    margin = list(l = 130),
    legend = list(x = 0.8, y = 0.1)
  )
```

### Green

```{r db_green}
pv_change %>% 
  filter(Party == "Green Party") %>% 
  mutate(Electorate = fct_reorder(Electorate, `2017`)) %>% 
  plot_ly(color = I("gray80")) %>% 
  add_segments(
    x = ~pmin(`2014`, `2017`), xend = ~pmax(`2014`, `2017`),
    y = ~Electorate, yend = ~Electorate,
    showlegend = FALSE
  ) %>% 
  add_markers(
    x = ~`2014`, y = ~Electorate, name = "2014",
    color = I("grey"), hoverinfo = "text",
    text = ~paste0(`2014`, "% of ", Electorate, " votes in 2014")
  ) %>% 
  add_markers(
    x = ~`2017`, y = ~Electorate, name = "2017",
    color = I(parties_v["Green"]), hoverinfo = "text",
    text = ~paste0(`2017`, "% of ", Electorate, " votes in 2017")
  ) %>% 
  layout(
    title = "Relative Votes for the Green Party by Electorate",
    xaxis = list(title = "Votes (Percent)"),
    yaxis = list(title = "", dtick = 1),
    margin = list(l = 130),
    legend = list(x = 0.8, y = 0.1)
  )
```

### Māori

```{r db_māori}
pv_change %>% 
  filter(Party == "Māori Party") %>% 
  mutate(Electorate = fct_reorder(Electorate, `2017`)) %>% 
  plot_ly(color = I("gray80")) %>% 
  add_segments(
    x = ~pmin(`2014`, `2017`), xend = ~pmax(`2014`, `2017`),
    y = ~Electorate, yend = ~Electorate,
    showlegend = FALSE
  ) %>% 
  add_markers(
    x = ~`2014`, y = ~Electorate, name = "2014",
    color = I("grey"), hoverinfo = "text",
    text = ~paste0(`2014`, "% of ", Electorate, " votes in 2014")
  ) %>% 
  add_markers(
    x = ~`2017`, y = ~Electorate, name = "2017",
    color = I(parties_v["Māori"]), hoverinfo = "text",
    text = ~paste0(`2017`, "% of ", Electorate, " votes in 2017")
  ) %>% 
  layout(
    title = "Relative Votes for the Māori Party by Electorate",
    xaxis = list(title = "Votes (Percent)"),
    yaxis = list(title = "", dtick = 1),
    margin = list(l = 130),
    legend = list(x = 0.8, y = 0.1)
  )
```

### ACT

```{r db_act}
pv_change %>% 
  filter(Party == "ACT New Zealand") %>% 
  mutate(Electorate = fct_reorder(Electorate, `2017`)) %>% 
  plot_ly(color = I("gray80")) %>% 
  add_segments(
    x = ~pmin(`2014`, `2017`), xend = ~pmax(`2014`, `2017`),
    y = ~Electorate, yend = ~Electorate,
    showlegend = FALSE
  ) %>% 
  add_markers(
    x = ~`2014`, y = ~Electorate, name = "2014",
    color = I("grey"), hoverinfo = "text",
    text = ~paste0(`2014`, "% of ", Electorate, " votes in 2014")
  ) %>% 
  add_markers(
    x = ~`2017`, y = ~Electorate, name = "2017",
    color = I(parties_v["ACT"]), hoverinfo = "text",
    text = ~paste0(`2017`, "% of ", Electorate, " votes in 2017")
  ) %>% 
  layout(
    title = "Relative Votes for the ACT Party by Electorate",
    xaxis = list(title = "Votes (Percent)"),
    yaxis = list(title = "", dtick = 1),
    margin = list(l = 130),
    legend = list(x = 0.8, y = 0.1)
  )
```

### Conservative

```{r db_conservative}
pv_change %>% 
  filter(Party == "Conservative") %>% 
  mutate(Electorate = fct_reorder(Electorate, `2017`)) %>% 
  plot_ly(color = I("gray80")) %>% 
  add_segments(
    x = ~pmin(`2014`, `2017`), xend = ~pmax(`2014`, `2017`),
    y = ~Electorate, yend = ~Electorate,
    showlegend = FALSE
  ) %>% 
  add_markers(
    x = ~`2014`, y = ~Electorate, name = "2014",
    color = I("grey"), hoverinfo = "text",
    text = ~paste0(`2014`, "% of ", Electorate, " votes in 2014")
  ) %>% 
  add_markers(
    x = ~`2017`, y = ~Electorate, name = "2017",
    color = I(parties_v["Conservative"]), hoverinfo = "text",
    text = ~paste0(`2017`, "% of ", Electorate, " votes in 2017")
  ) %>% 
  layout(
    title = "Relative Votes for the Conservative Party by Electorate",
    xaxis = list(title = "Votes (Percent)"),
    yaxis = list(title = "", dtick = 1),
    margin = list(l = 130),
    legend = list(x = 0.8, y = 0.1)
  )
```

### Left vs Right

```{r db_left_right}
pv_change_lr <-
  pv_change %>% 
  mutate(Group = 
           case_when(Party %in% c("Labour Party", "Green Party") ~ "Left",
                     Party %in% c("National Party", "ACT New Zealand", "Conservative") ~ "Right")) %>%
  filter(!is.na(Group)) %>%
  group_by(Group, Electorate) %>%
  summarise(`2014` = sum(`2014`), `2017` = sum(`2017`)) %>%
  ungroup()
pv_change_lr <-
  inner_join(
    pv_change_lr %>% 
      filter(Group == "Left") %>% 
      select(Electorate, `Left 2014` = `2014`, `Left 2017` = `2017`),
    pv_change_lr %>% 
      filter(Group == "Right") %>% 
      select(Electorate, `Right 2014` = `2014`, `Right 2017` = `2017`)
  ) %>% 
  mutate(Electorate = fct_reorder(Electorate, `Left 2017`))
plot_ly(pv_change_lr, color = I("gray80")) %>% 
  add_segments(
    x = ~pmin(`Left 2014`, `Left 2017`), xend = ~pmax(`Left 2014`, `Left 2017`),
    y = ~Electorate, yend = ~Electorate,
    showlegend = FALSE
  ) %>% 
  add_segments(
    x = ~pmin(`Right 2014`, `Right 2017`), xend = ~pmax(`Right 2014`, `Right 2017`),
    y = ~Electorate, yend = ~Electorate,
    showlegend = FALSE
  ) %>% 
  add_markers(
    x = ~`Left 2014`, y = ~Electorate, name = "Left 2014",
    color = I("gray80"), hoverinfo = "text",
    text = ~paste0(`Left 2014`, "% of ", Electorate, " voted for Lab/Grn in 2014")
  ) %>% 
  add_markers(
    x = ~`Right 2014`, y = ~Electorate, name = "Right 2014",
    color = I("gray80"), hoverinfo = "text",
    text = ~paste0(`Right 2014`, "% of ", Electorate, " voted for Nat/ACT/Cons in 2014")
  ) %>% 
  add_markers(
    x = ~`Left 2017`, y = ~Electorate, name = "Left 2017",
    color = I("red1"), hoverinfo = "text",
    text = ~paste0(`Left 2017`, "% of ", Electorate, " voted for Lab/Grn in 2017")
  ) %>% 
  add_markers(
    x = ~`Right 2017`, y = ~Electorate, name = "Right 2017",
    color = I("blue1"), hoverinfo = "text",
    text = ~paste0(`Right 2017`, "% of ", Electorate, " voted for Nat/ACT/Cons in 2017")
  ) %>% 
  layout(
    title = "Left and Right Votes by Electorate",
    xaxis = list(title = "Votes (Percent)", dtick = 10),
    yaxis = list(title = "", dtick = 1),
    margin = list(l = 130)
  )
```




## Explanation {.sidebar}

These dumbbell plots show the percent of the party vote received in each electorate,
highlighting the shift from 2014.

National had mostly decreases, with a few small increases.

The Greens had universal decreases, and Labour had corresponding universal increases.

New Zealand First mostly had decreases, some quite large. Apart from the Winston effect
in Northland, the most notable increase is in Clutha-Southland, scene of recent controversy.

The Māori Party decreased in six of the Māori seats, but had small gains in South 
Auckland. ACT pretty much lost all its "large" (1.5-3%) support. And the Conservatives
lost all their support!

The final one compares votes for "the Left" (Labour and Green) vs "the Right" (National,
ACT and Conservative). Apart from Manukau East, there's been a universal shift left.





<script>
  (function(i,s,o,g,r,a,m){i['GoogleAnalyticsObject']=r;i[r]=i[r]||function(){
  (i[r].q=i[r].q||[]).push(arguments)},i[r].l=1*new Date();a=s.createElement(o),
  m=s.getElementsByTagName(o)[0];a.async=1;a.src=g;m.parentNode.insertBefore(a,m)
  })(window,document,'script','https://www.google-analytics.com/analytics.js','ga');

  ga('create', 'UA-30013353-2', 'auto');
  ga('send', 'pageview');

</script>
