---
title: "2017 New Zealand Election (Provisional Results)"
output: 
  flexdashboard::flex_dashboard:
    theme: spacelab
    self_contained: false
    lib_dir: lib
    mathjax: null
    navbar:
      - { title: "@dakvid", icon: "fa-twitter", href: "https://twitter.com/dakvid", align: right}
    social: menu
    source_code: https://github.com/dakvid/election2017
---

```{r init, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```

```{r setup, include=FALSE}
library(flexdashboard)
library(sf)
library(magrittr)
library(dplyr)
library(forcats)
library(purrr)
library(stringr)
library(leaflet)
library(leaflet.minicharts)
library(nzelect)
library(htmltools)
library(htmlwidgets)
library(plotly)

begin_label_text <- "<div style='font-size:12px;float:left'>"
begin_label_text_200 <- "<div style='font-size:12px;width:200px;float:left'>"
begin_label_header <- "<span style='font-size:18px;font-weight:bold'>"
end_label_header <- "</span><br>"
end_label_text <- "</div>"

load("data_geo/nzhex.rda")
load("data_votes/preliminary/GE2017_prelim.rda")
data("GE2014")
GE2014 %<>%
  mutate(id = Electorate %>% str_extract("[0-9]+") %>% as.integer(),
         Party = if_else(Party == "Maori Party", "Māori Party", Party))

party_voters_2017 <-
  GE2017_prelim %>% 
  filter(VotingType == "Party",
         !str_detect(Party, "nformal")) %>% 
  group_by(id) %>% 
  summarise(Voters = sum(Votes)) %>% 
  ungroup()

candidate_voters_2017 <-
  GE2017_prelim %>% 
  filter(VotingType == "Candidate",
         !str_detect(Candidate, "nformal")) %>% 
  group_by(id) %>% 
  summarise(Voters = sum(Votes)) %>% 
  ungroup()

party_voters_2014 <-
  GE2014 %>% 
  filter(VotingType == "Party",
         !str_detect(Party, "nformal")) %>% 
  group_by(id) %>% 
  summarise(Voters2014 = sum(Votes)) %>% 
  ungroup()

candidate_voters_2014 <-
  GE2014 %>% 
  filter(VotingType == "Candidate",
         !str_detect(Candidate, "nformal")) %>% 
  group_by(id) %>% 
  summarise(Voters2014 = sum(Votes)) %>% 
  ungroup()

getLeafletOptions <- function(minZoom, maxZoom, ...) {
  leafletOptions(
    crs = leafletCRS("L.CRS.Simple"),
    minZoom = minZoom, maxZoom = maxZoom,
    dragging = TRUE, zoomControl = FALSE,
    tap = FALSE,
    attributionControl = FALSE , ...)
}

getFactorPal <- function(f) {
  colorFactor(colormap::colormap(
    colormap = colormap::colormaps$hsv,
    nshades = length(f)), f)
}
```

# Hexmaps {.storyboard data-icon="fa-map"}

### Hello HexMaps: a hexagonal cartogram of New Zealand electorates (2014-2017)

```{r hex_regions}
leaflet(options = getLeafletOptions(5.8, 5.8)) %>% 
  addPolygons(
    data = nzhex,
    label = ~nzhex %>% 
      transpose() %>% 
      map(~ paste0(begin_label_text,
               begin_label_header, .x$electorate_name, end_label_header,
               "In ", .x$region, ".",
               end_label_text) %>% HTML()),
    weight=2, color='#000000', group = 'Regions',
    fillOpacity = 0.6, opacity = 1, fillColor = ~getFactorPal(region)(region),
    highlightOptions = highlightOptions(weight = 4)
  ) %>% 
  addLabelOnlyMarkers(
    data = nzhex %>% st_centroid(),
    label = ~abbr,
    labelOptions = labelOptions(
      noHide = 'T', textOnly = T,
      offset=c(-6,-11))
  ) %>% 
  onRender(
    "function(el, t) {
      var myMap = this;
      // get rid of the ugly grey background
      myMap._container.style['background'] = '#ffffff';
      myMap.scrollWheelZoom.disable();
    }"
  )
```

***

This is a [cartogram](https://en.wikipedia.org/wiki/Cartogram) of the current
(2014-2017) New Zealand electorates, with one hexagon for each electorate.
Electorates have roughly equal populations, so this is a reasonable map of
population distribution. The North and South Islands should be obvious; the
Māori electorates make up a second mini NZ in the bottom right.

Why are these maps more useful than a geographic map? In short, because they don't
overrepresent rural areas and underrepresent urban areas.
[Stephan Beban (NZ)](https://thespinoff.co.nz/politics/24-09-2017/a-better-visual-breakdown-of-the-2017-election-results/) and
[Pitch Interactive (USA)](http://pitchinteractive.com/latest/tilegrams-more-human-maps/)
have more to say on the topic.

The colours here indicate the regional groupings, to help orient you to the
distorted geography. Many electorates cross regional boundaries, so these
are just a rough guide. Most notably, the Wellington region is split between
several Māori electorates. (Credit for the regional groupings to
[Chris Knox in the Herald](http://insights.nzherald.co.nz/election/)).

This hexagonal layout is due to 
[Joseph Wright](http://mapdruid.blogspot.co.nz/2015/05/hexagonal-tile-map-of-new-zealand.html),
and I used his shapefiles as a base.

Chris McDowall made [a different hexagonal layout](http://hindsight.clerestories.com/2014/01/06/chris-mcdowall-hexagonal-maps/)
for the 2011 electorates, which [I adapted](http://david.frigge.nz/nzhex2014/) in 2014.
I think I like that better in terms of the geographic layout, but it is very
skinny so it's less compact for on-screen display.

Click through the storyboard and see different data from the election
visualised in hexmaps...


### Electorate Winners: Who won the seats?

```{r hex_candidates}
electorate_winners <-
  GE2017_prelim %>%
  filter(VotingType == "Candidate") %>%
  group_by(id, Party) %>%
  summarise(Votes = sum(Votes)) %>%
  top_n(1, Votes) %>%
  ungroup() %>% 
  inner_join(candidate_voters_2017) %>% 
  mutate(PerCent = round(Votes / Voters * 100, 1)) %>% 
  mutate(party_colour = case_when(
    Party == "National Party" ~ parties_v["National"],
    Party == "Labour Party" ~ parties_v["Labour"],
    Party == "New Zealand First Party" ~ parties_v["NZ First"],
    Party == "ACT New Zealand" ~ parties_v["ACT"],
    Party == "United Future" ~ parties_v["United Future"],
    Party == "Māori Party" ~ parties_v["Māori"]
  )) %>% 
  mutate(opa = 0.1 + 0.85 * ((PerCent - min(PerCent)) / (max(PerCent) - min(PerCent))))
electorate_winners$labeltext <-
  electorate_winners %>% 
  inner_join(nzhex) %>% 
  transpose() %>% 
  map(~ paste0(begin_label_text,
               begin_label_header, .x$electorate_name, end_label_header,
               .x$Party, " won with ", prettyNum(.x$Votes, big.mark=','), " votes.",
               "<br>",
               "<div style='width:95%'>",
               "<span style='color:", .x$party_colour, ";float:left'>", .x$PerCent, "%</span>",
               "<br clear='all' />",
               "<span style='background:", .x$party_colour, ";width:", .x$PerCent, "%;float:left'>&nbsp;</span>",
               "</div>",
               end_label_text) %>% 
        HTML())

leaflet(options = getLeafletOptions(5.8, 5.8)) %>% 
  addPolygons(
    data = inner_join(nzhex, electorate_winners),
    label = ~labeltext,
    weight=2, color='#000000', group = 'Electorate Seats',
    fillOpacity = ~opa, opacity = 1, fillColor = ~party_colour,
    highlightOptions = highlightOptions(weight = 4)
  ) %>% 
  addLabelOnlyMarkers(
    data = nzhex %>% st_centroid(),
    label = ~abbr,
    labelOptions = labelOptions(
      noHide = 'T', textOnly = T,
      offset=c(-6,-11))
  ) %>% 
  onRender(
    "function(el, t) {
      var myMap = this;
      // get rid of the ugly grey background
      myMap._container.style['background'] = '#ffffff';
    }"
  )
```

***

Each hexagon is coloured for the party of the winning candidate. The strength of colour tracks
the percentage of votes the winner received --- from
`r electorate_winners %>% top_n(-1, PerCent) %$% PerCent`% for the
`r electorate_winners %>% top_n(-1, PerCent) %$% Party` in
`r electorate_winners %>% top_n(-1, PerCent) %>% inner_join(nzhex) %$% electorate_name`
to
`r electorate_winners %>% top_n(1, PerCent) %$% PerCent`% for the
`r electorate_winners %>% top_n(1, PerCent) %$% Party` in
`r electorate_winners %>% top_n(1, PerCent) %>% inner_join(nzhex) %$% electorate_name`.


### Party Votes: How are each party's voters distributed? Do they have broad or localised support?

```{r hex_parties}
party_votes <-
  GE2017_prelim %>% 
  filter(VotingType == "Party") %>% 
  group_by(id, Party) %>% 
  summarise(Votes = sum(Votes)) %>% 
  ungroup() %>% 
  inner_join(party_voters_2017) %>% 
  mutate(PerCent = round(Votes / Voters * 100, 1)) %>% 
  mutate(party_colour = case_when(
    Party == "National Party" ~ parties_v["National"],
    Party == "Labour Party" ~ parties_v["Labour"],
    Party == "New Zealand First Party" ~ parties_v["NZ First"],
    Party == "ACT New Zealand" ~ parties_v["ACT"],
    Party == "United Future" ~ parties_v["United Future"],
    Party == "Māori Party" ~ parties_v["Māori"],
    Party == "Green Party" ~ parties_v["Green"],
    Party == "The Opportunities Party (TOP)" ~ parties_v["TOP"],
    Party == "Aotearoa Legalise Cannabis Party" ~ parties_v["Legalise Cannabis"],
    Party == "Ban1080" ~ parties_v["Ban 1080"],
    Party == "Conservative" ~ parties_v["Conservative"],
    Party == "Internet Party" ~ parties_v["Internet"],
    Party == "MANA" ~ parties_v["Mana"]
  )) %>% 
  group_by(Party) %>% 
  mutate(mean_votes = mean(Votes),
         opa = case_when(
           Votes < 0.8 * mean_votes ~ 0.15,
           Votes < 0.95 * mean_votes ~ 0.3,
           Votes <= 1.05 * mean_votes ~ 0.5,
           Votes <= 1.2 * mean_votes ~ 0.7,
           Votes > 1.2 * mean_votes ~ 0.85
         ),
         MaxPC = ceiling(max(PerCent))
  ) %>% 
  ungroup()

party_votes_2014 <-
  GE2014 %>% 
  filter(VotingType == "Party") %>% 
  group_by(id, Party) %>% 
  summarise(Votes2014 = sum(Votes)) %>% 
  ungroup() %>% 
  inner_join(party_voters_2014) %>% 
  mutate(PerCent2014 = round(Votes2014 / Voters2014 * 100, 1)) %>% 
  group_by(Party) %>% 
  mutate(MaxPC2014 = ceiling(max(PerCent2014))) %>% 
  ungroup()

party_votes$labeltext <-
  party_votes %>% 
  left_join(party_votes_2014) %>% 
  group_by(Party) %>% 
  mutate(MaxPC = if_else(is.na(MaxPC2014), MaxPC, max(MaxPC, MaxPC2014))) %>% 
  ungroup() %>% 
  inner_join(nzhex) %>% 
  transpose() %>% 
  map(~ paste0(begin_label_text_200,
               begin_label_header, .x$electorate_name, end_label_header,
               prettyNum(.x$Votes, big.mark=','), " votes.",
               "<br>",
               "<div style='width:95%'>",
               "<span style='float:left'>2017: ", .x$PerCent, "%</span>",
               "<br clear='all' />",
               "<span style='background:", .x$party_colour, ";width:", round(.x$PerCent / .x$MaxPC * 100), "%;float:left'>&nbsp;</span>",
               if (is.na(.x$MaxPC2014)) {""} else {
                 paste0(
                   "<br clear='all' />",
                   "<span style='float:left'>2014: ", .x$PerCent2014, "%</span>",
                   "<br clear='all' />",
                   "<span style='background:", .x$party_colour, ";width:", round(.x$PerCent2014 / .x$MaxPC * 100), "%;float:left'>&nbsp;</span>"
                 )
               },
               "</div>",
               end_label_text) %>% 
        HTML())

leaflet(options = getLeafletOptions(5.8, 5.8)) %>% 
  addPolygons(
    data = nzhex %>% 
      inner_join(party_votes %>% filter(Party == "National Party")),
    label = ~labeltext,
    weight=2, color='#000000', group = 'National 46.0%',
    fillOpacity = ~opa, opacity = 1, fillColor = ~party_colour,
    highlightOptions = highlightOptions(weight = 4)
  ) %>% 
  addPolygons(
    data = nzhex %>% 
      inner_join(party_votes %>% filter(Party == "Labour Party")),
    label = ~labeltext,
    weight=2, color='#000000', group = 'Labour 35.8%',
    fillOpacity = ~opa, opacity = 1, fillColor = ~party_colour,
    highlightOptions = highlightOptions(weight = 4)
  ) %>% 
  addPolygons(
    data = nzhex %>% 
      inner_join(party_votes %>% filter(Party == "New Zealand First Party")),
    label = ~labeltext,
    weight=2, color='#000000', group = 'NZ First 7.5%',
    fillOpacity = ~opa, opacity = 1, fillColor = ~party_colour,
    highlightOptions = highlightOptions(weight = 4)
  ) %>% 
  addPolygons(
    data = nzhex %>% 
      inner_join(party_votes %>% filter(Party == "Green Party")),
    label = ~labeltext,
    weight=2, color='#000000', group = 'Green 5.9%',
    fillOpacity = ~opa, opacity = 1, fillColor = ~party_colour,
    highlightOptions = highlightOptions(weight = 4)
  ) %>% 
  addPolygons(
    data = nzhex %>% 
      inner_join(party_votes %>% filter(Party == "The Opportunities Party (TOP)")),
    label = ~labeltext,
    weight=2, color='#000000', group = 'TOP 2.2%',
    fillOpacity = ~opa, opacity = 1, fillColor = ~party_colour,
    highlightOptions = highlightOptions(weight = 4)
  ) %>% 
  addPolygons(
    data = nzhex %>% 
      inner_join(party_votes %>% filter(Party == "Māori Party")),
    label = ~labeltext,
    weight=2, color='#000000', group = 'Māori 1.1%',
    fillOpacity = ~opa, opacity = 1, fillColor = ~party_colour,
    highlightOptions = highlightOptions(weight = 4)
  ) %>% 
  addPolygons(
    data = nzhex %>% 
      inner_join(party_votes %>% filter(Party == "ACT New Zealand")),
    label = ~labeltext,
    weight=2, color='#000000', group = 'ACT 0.5%',
    fillOpacity = ~opa, opacity = 1, fillColor = ~party_colour,
    highlightOptions = highlightOptions(weight = 4)
  ) %>% 
  addPolygons(
    data = nzhex %>% 
      inner_join(party_votes %>% filter(Party == "Aotearoa Legalise Cannabis Party")),
    label = ~labeltext,
    weight=2, color='#000000', group = 'Cannabis 0.3%',
    fillOpacity = ~opa, opacity = 1, fillColor = ~party_colour,
    highlightOptions = highlightOptions(weight = 4)
  ) %>% 
  addPolygons(
    data = nzhex %>% 
      inner_join(party_votes %>% filter(Party == "Conservative")),
    label = ~labeltext,
    weight=2, color='#000000', group = 'Conservative 0.2%',
    fillOpacity = ~opa, opacity = 1, fillColor = ~party_colour,
    highlightOptions = highlightOptions(weight = 4)
  ) %>% 
  addLabelOnlyMarkers(
    data = nzhex %>% st_centroid(),
    label = ~abbr,
    labelOptions = labelOptions(
      noHide = 'T', textOnly = T,
      offset=c(-6,-11))
  ) %>% 
  addLayersControl(
    baseGroups = c("National 46.0%", "Labour 35.8%", "NZ First 7.5%", "Green 5.9%",
                   "TOP 2.2%", "Māori 1.1%", "ACT 0.5%", "Cannabis 0.3%", "Conservative 0.2%"),
    options = layersControlOptions(collapsed = FALSE)
  ) %>% 
  onRender(
    "function(el, t) {
      var myMap = this;
      // get rid of the ugly grey background
      myMap._container.style['background'] = '#ffffff';
    }"
  )
```

***

Choose a party and see the tiles shaded relative to the number of voters. There are
five shades: within 5% of the average, between 5-20% over (or under) average, and more than 20% over (or under) the average.
The mouseover compares the 2017 and 2014 percentages.

Note that because it's based on absolute numbers rather than percentages, the Māori
electorates may be unexpectedly lighter because of their lower turnouts.

National and Labour have fairly complementary patterns, as do NZ First and the Greens.

The collapse of the Conservative vote and the redistribution of the 4% wasted votes in
2014 is a factor in the 2017 results that often gets overlooked.


### Coalition votes: Which electorates were "won" by the incumbent government?

```{r hex_coalition}
coalition_votes <-
  party_votes %>% 
  filter(Party %in% c("National Party", "United Future", "Māori Party", "ACT New Zealand")) %>% 
  group_by(id) %>% 
  summarise(Votes = sum(Votes),
            Voters = min(Voters),
            party_colour = min(party_colour)) %>% # min=National
  mutate(PerCent = round(Votes / Voters * 100, 1),
         MaxPC = ceiling(max(PerCent))) %>% 
  ungroup() %>% 
  mutate(win = PerCent > 50)

coalition_votes_2014 <-
  party_votes_2014 %>% 
  filter(Party %in% c("National Party", "United Future", "Māori Party", "ACT New Zealand")) %>% 
  group_by(id) %>% 
  summarise(Votes2014 = sum(Votes2014),
            Voters2014 = min(Voters2014)) %>%
  mutate(PerCent2014 = round(Votes2014 / Voters2014 * 100, 1),
         MaxPC2014 = ceiling(max(PerCent2014))) %>% 
  ungroup() %>% 
  mutate(win2014 = PerCent2014 > 50)

coalition_votes$labeltext <-
  coalition_votes %>% 
  left_join(coalition_votes_2014) %>% 
  inner_join(nzhex) %>% 
  transpose() %>% 
  map(~ paste0(begin_label_text_200,
               begin_label_header, .x$electorate_name, end_label_header,
               prettyNum(.x$Votes, big.mark=','), " votes.",
               "<br>",
               "<div style='width:95%'>",
               "<span style='float:left'>2017: ", .x$PerCent, "%</span>",
               "<br clear='all' />",
               "<span style='background:", .x$party_colour, ";width:", round(.x$PerCent / .x$MaxPC * 100), "%;float:left'>&nbsp;</span>",
               if (is.na(.x$MaxPC2014)) {""} else {
                 paste0(
                   "<br clear='all' />",
                   "<span style='float:left'>2014: ", .x$PerCent2014, "%</span>",
                   "<br clear='all' />",
                   "<span style='background:", .x$party_colour, ";width:", round(.x$PerCent2014 / .x$MaxPC * 100), "%;float:left'>&nbsp;</span>"
                 )
               },
               "</div>",
               end_label_text) %>% 
        HTML())

leaflet(options = getLeafletOptions(5.8, 5.8)) %>% 
  addMinicharts(
    lng = nzhex %>% 
      arrange(id) %>% 
      st_centroid() %>% 
      st_coordinates() %>% 
      extract(, "X"),
    lat = nzhex %>% 
      arrange(id) %>% 
      st_centroid() %>% 
      st_coordinates() %>% 
      extract(, "Y"),
    chartdata = coalition_votes %>% 
      arrange(id) %>% 
      transmute(Coalition = PerCent, `"Change"` = 100 - PerCent),
    type = "pie",
    colorPalette = c("darkblue", "orange")
  ) %>%
  addPolygons(
    data = inner_join(nzhex, coalition_votes),
    label = ~labeltext,
    weight=2, color='#000000', group = 'Coalition Voters',
    fillColor = "black", fillOpacity = ~if_else(win, 0.2, 0),
    highlightOptions = highlightOptions(weight = 4)
  ) %>%
  addLabelOnlyMarkers(
    data = nzhex %>% st_centroid(),
    label = ~abbr,
    labelOptions = labelOptions(
      style = list("color" = "white"),
      noHide = 'T', textOnly = T,
      offset=c(-6,-11))
  ) %>%
  onRender(
    "function(el, t) {
      var myMap = this;
      // get rid of the ugly grey background
      myMap._container.style['background'] = '#ffffff';
    }"
  )
```

***

The incumbant government is a four-party coalition between National (47.0% in 2014),
Māori (1.3%), ACT (0.7%) and United Future (0.2%). Some have argued that the election
was a referendum on the government and that the majority voted against it.

Though I'm no fan of pie charts in general, I think they're reasonable here to show the
proportion of votes for and against coalition parties, with the latter arguably in
favour of some sort of change.

The coalition received a majority in just 
`r sum(coalition_votes$win)` of 71 electorates, down from
`r sum(coalition_votes_2014$win2014)` in 2014.

They increased their vote in
`r inner_join(coalition_votes, coalition_votes_2014) %>% filter(PerCent > PerCent2014) %>% nrow()`
electorates, but none by more than 4% (the 2014 national vote for the collapsed
Conservative party).
The decreases ranged up to
`r inner_join(coalition_votes, coalition_votes_2014) %>% inner_join(nzhex) %>% arrange(PerCent2014 - PerCent, PerCent) %>% tail(1) %$% {PerCent2014 - PerCent}`%
in
`r inner_join(coalition_votes, coalition_votes_2014) %>% inner_join(nzhex) %>% arrange(PerCent2014 - PerCent, PerCent) %>% tail(1) %$% electorate_name`.



# Dumbbells {data-icon="ion-ios-settings-strong"}

```{r db_data}
pv_change <-
  inner_join(
    party_votes %>% 
      inner_join(nzhex) %>% 
      select(Party, Electorate = electorate_name, `2017` = PerCent),
    party_votes_2014 %>% 
      inner_join(nzhex) %>% 
      select(Party, Electorate = electorate_name, `2014` = PerCent2014)
  )
```


## Plots {.tabset}

### National

```{r db_national}
pv_change %>% 
  filter(Party == "National Party") %>% 
  mutate(Electorate = fct_reorder(Electorate, `2017`)) %>% 
  plot_ly(color = I("gray80")) %>% 
  add_segments(
    x = ~pmin(`2014`, `2017`), xend = ~pmax(`2014`, `2017`),
    y = ~Electorate, yend = ~Electorate,
    showlegend = FALSE
  ) %>% 
  add_markers(
    x = ~`2014`, y = ~Electorate, name = "2014",
    color = I("grey"), hoverinfo = "text",
    text = ~paste0(`2014`, "% of ", Electorate, " votes in 2014")
  ) %>% 
  add_markers(
    x = ~`2017`, y = ~Electorate, name = "2017",
    color = I(parties_v["National"]), hoverinfo = "text",
    text = ~paste0(`2017`, "% of ", Electorate, " votes in 2017")
  ) %>% 
  layout(
    title = "Relative Votes for the National Party by Electorate",
    xaxis = list(title = "Votes (Percent)"),
    yaxis = list(title = "", dtick = 1),
    margin = list(l = 130),
    legend = list(x = 0.8, y = 0.1)
  )
```

### Labour

```{r db_labour}
pv_change %>% 
  filter(Party == "Labour Party") %>% 
  mutate(Electorate = fct_reorder(Electorate, `2017`)) %>% 
  plot_ly(color = I("gray80")) %>% 
  add_segments(
    x = ~pmin(`2014`, `2017`), xend = ~pmax(`2014`, `2017`),
    y = ~Electorate, yend = ~Electorate,
    showlegend = FALSE
  ) %>% 
  add_markers(
    x = ~`2014`, y = ~Electorate, name = "2014",
    color = I("grey"), hoverinfo = "text",
    text = ~paste0(`2014`, "% of ", Electorate, " votes in 2014")
  ) %>% 
  add_markers(
    x = ~`2017`, y = ~Electorate, name = "2017",
    color = I(parties_v["Labour"]), hoverinfo = "text",
    text = ~paste0(`2017`, "% of ", Electorate, " votes in 2017")
  ) %>% 
  layout(
    title = "Relative Votes for the Labour Party by Electorate",
    xaxis = list(title = "Votes (Percent)"),
    yaxis = list(title = "", dtick = 1),
    margin = list(l = 130),
    legend = list(x = 0.8, y = 0.1)
  )
```

### NZ First

```{r db_nzf}
pv_change %>% 
  filter(Party == "New Zealand First Party") %>% 
  mutate(Electorate = fct_reorder(Electorate, `2017`)) %>% 
  plot_ly(color = I("gray80")) %>% 
  add_segments(
    x = ~pmin(`2014`, `2017`), xend = ~pmax(`2014`, `2017`),
    y = ~Electorate, yend = ~Electorate,
    showlegend = FALSE
  ) %>% 
  add_markers(
    x = ~`2014`, y = ~Electorate, name = "2014",
    color = I("grey"), hoverinfo = "text",
    text = ~paste0(`2014`, "% of ", Electorate, " votes in 2014")
  ) %>% 
  add_markers(
    x = ~`2017`, y = ~Electorate, name = "2017",
    color = I(parties_v["NZ First"]), hoverinfo = "text",
    text = ~paste0(`2017`, "% of ", Electorate, " votes in 2017")
  ) %>% 
  layout(
    title = "Relative Votes for the NZ First Party by Electorate",
    xaxis = list(title = "Votes (Percent)"),
    yaxis = list(title = "", dtick = 1),
    margin = list(l = 130),
    legend = list(x = 0.8, y = 0.1)
  )
```

### Green

```{r db_green}
pv_change %>% 
  filter(Party == "Green Party") %>% 
  mutate(Electorate = fct_reorder(Electorate, `2017`)) %>% 
  plot_ly(color = I("gray80")) %>% 
  add_segments(
    x = ~pmin(`2014`, `2017`), xend = ~pmax(`2014`, `2017`),
    y = ~Electorate, yend = ~Electorate,
    showlegend = FALSE
  ) %>% 
  add_markers(
    x = ~`2014`, y = ~Electorate, name = "2014",
    color = I("grey"), hoverinfo = "text",
    text = ~paste0(`2014`, "% of ", Electorate, " votes in 2014")
  ) %>% 
  add_markers(
    x = ~`2017`, y = ~Electorate, name = "2017",
    color = I(parties_v["Green"]), hoverinfo = "text",
    text = ~paste0(`2017`, "% of ", Electorate, " votes in 2017")
  ) %>% 
  layout(
    title = "Relative Votes for the Green Party by Electorate",
    xaxis = list(title = "Votes (Percent)"),
    yaxis = list(title = "", dtick = 1),
    margin = list(l = 130),
    legend = list(x = 0.8, y = 0.1)
  )
```

### Māori

```{r db_māori}
pv_change %>% 
  filter(Party == "Māori Party") %>% 
  mutate(Electorate = fct_reorder(Electorate, `2017`)) %>% 
  plot_ly(color = I("gray80")) %>% 
  add_segments(
    x = ~pmin(`2014`, `2017`), xend = ~pmax(`2014`, `2017`),
    y = ~Electorate, yend = ~Electorate,
    showlegend = FALSE
  ) %>% 
  add_markers(
    x = ~`2014`, y = ~Electorate, name = "2014",
    color = I("grey"), hoverinfo = "text",
    text = ~paste0(`2014`, "% of ", Electorate, " votes in 2014")
  ) %>% 
  add_markers(
    x = ~`2017`, y = ~Electorate, name = "2017",
    color = I(parties_v["Māori"]), hoverinfo = "text",
    text = ~paste0(`2017`, "% of ", Electorate, " votes in 2017")
  ) %>% 
  layout(
    title = "Relative Votes for the Māori Party by Electorate",
    xaxis = list(title = "Votes (Percent)"),
    yaxis = list(title = "", dtick = 1),
    margin = list(l = 130),
    legend = list(x = 0.8, y = 0.1)
  )
```

### ACT

```{r db_act}
pv_change %>% 
  filter(Party == "ACT New Zealand") %>% 
  mutate(Electorate = fct_reorder(Electorate, `2017`)) %>% 
  plot_ly(color = I("gray80")) %>% 
  add_segments(
    x = ~pmin(`2014`, `2017`), xend = ~pmax(`2014`, `2017`),
    y = ~Electorate, yend = ~Electorate,
    showlegend = FALSE
  ) %>% 
  add_markers(
    x = ~`2014`, y = ~Electorate, name = "2014",
    color = I("grey"), hoverinfo = "text",
    text = ~paste0(`2014`, "% of ", Electorate, " votes in 2014")
  ) %>% 
  add_markers(
    x = ~`2017`, y = ~Electorate, name = "2017",
    color = I(parties_v["ACT"]), hoverinfo = "text",
    text = ~paste0(`2017`, "% of ", Electorate, " votes in 2017")
  ) %>% 
  layout(
    title = "Relative Votes for the ACT Party by Electorate",
    xaxis = list(title = "Votes (Percent)"),
    yaxis = list(title = "", dtick = 1),
    margin = list(l = 130),
    legend = list(x = 0.8, y = 0.1)
  )
```

### Conservative

```{r db_conservative}
pv_change %>% 
  filter(Party == "Conservative") %>% 
  mutate(Electorate = fct_reorder(Electorate, `2017`)) %>% 
  plot_ly(color = I("gray80")) %>% 
  add_segments(
    x = ~pmin(`2014`, `2017`), xend = ~pmax(`2014`, `2017`),
    y = ~Electorate, yend = ~Electorate,
    showlegend = FALSE
  ) %>% 
  add_markers(
    x = ~`2014`, y = ~Electorate, name = "2014",
    color = I("grey"), hoverinfo = "text",
    text = ~paste0(`2014`, "% of ", Electorate, " votes in 2014")
  ) %>% 
  add_markers(
    x = ~`2017`, y = ~Electorate, name = "2017",
    color = I(parties_v["Conservative"]), hoverinfo = "text",
    text = ~paste0(`2017`, "% of ", Electorate, " votes in 2017")
  ) %>% 
  layout(
    title = "Relative Votes for the Conservative Party by Electorate",
    xaxis = list(title = "Votes (Percent)"),
    yaxis = list(title = "", dtick = 1),
    margin = list(l = 130),
    legend = list(x = 0.8, y = 0.1)
  )
```

### Left vs Right

```{r db_left_right}
pv_change_lr <-
  pv_change %>% 
  mutate(Group = 
           case_when(Party %in% c("Labour Party", "Green Party") ~ "Left",
                     Party %in% c("National Party", "ACT New Zealand", "Conservative") ~ "Right")) %>%
  filter(!is.na(Group)) %>%
  group_by(Group, Electorate) %>%
  summarise(`2014` = sum(`2014`), `2017` = sum(`2017`)) %>%
  ungroup()
pv_change_lr <-
  inner_join(
    pv_change_lr %>% 
      filter(Group == "Left") %>% 
      select(Electorate, `Left 2014` = `2014`, `Left 2017` = `2017`),
    pv_change_lr %>% 
      filter(Group == "Right") %>% 
      select(Electorate, `Right 2014` = `2014`, `Right 2017` = `2017`)
  ) %>% 
  mutate(Electorate = fct_reorder(Electorate, `Left 2017`))
plot_ly(pv_change_lr, color = I("gray80")) %>% 
  add_segments(
    x = ~pmin(`Left 2014`, `Left 2017`), xend = ~pmax(`Left 2014`, `Left 2017`),
    y = ~Electorate, yend = ~Electorate,
    showlegend = FALSE
  ) %>% 
  add_segments(
    x = ~pmin(`Right 2014`, `Right 2017`), xend = ~pmax(`Right 2014`, `Right 2017`),
    y = ~Electorate, yend = ~Electorate,
    showlegend = FALSE
  ) %>% 
  add_markers(
    x = ~`Left 2014`, y = ~Electorate, name = "Left 2014",
    color = I("gray80"), hoverinfo = "text",
    text = ~paste0(`Left 2014`, "% of ", Electorate, " voted for Lab/Grn in 2014")
  ) %>% 
  add_markers(
    x = ~`Right 2014`, y = ~Electorate, name = "Right 2014",
    color = I("gray80"), hoverinfo = "text",
    text = ~paste0(`Right 2014`, "% of ", Electorate, " voted for Nat/ACT/Cons in 2014")
  ) %>% 
  add_markers(
    x = ~`Left 2017`, y = ~Electorate, name = "Left 2017",
    color = I("red1"), hoverinfo = "text",
    text = ~paste0(`Left 2017`, "% of ", Electorate, " voted for Lab/Grn in 2017")
  ) %>% 
  add_markers(
    x = ~`Right 2017`, y = ~Electorate, name = "Right 2017",
    color = I("blue1"), hoverinfo = "text",
    text = ~paste0(`Right 2017`, "% of ", Electorate, " voted for Nat/ACT/Cons in 2017")
  ) %>% 
  layout(
    title = "Left and Right Votes by Electorate",
    xaxis = list(title = "Votes (Percent)", dtick = 10),
    yaxis = list(title = "", dtick = 1),
    margin = list(l = 130)
  )
```




## Explanation {.sidebar}

These dumbbell plots show the percent of the party vote received in each electorate,
showing the shift from 2014.

National had mostly decreases, with a few small increases.

The Greens had universal decreases, and Labour had corresponding universal increases.

New Zealand First mostly had decreases, some quite large. Apart from the Winston effect
in Northland, the most notable increase is in Clutha-Southland, scene of recent controversy.

The Māori Party decreased in six of the Māori seats, but had small gains in South 
Auckland. ACT pretty much lost all its "large" (1.5-3%) support. And the Conservatives
lost all their support!

The final one compares votes for "the Left" (Labour and Green) vs "the Right" (National,
ACT and Conservative). Apart from Manukau East, there's been a universal shift left.


<script>
  (function(i,s,o,g,r,a,m){i['GoogleAnalyticsObject']=r;i[r]=i[r]||function(){
  (i[r].q=i[r].q||[]).push(arguments)},i[r].l=1*new Date();a=s.createElement(o),
  m=s.getElementsByTagName(o)[0];a.async=1;a.src=g;m.parentNode.insertBefore(a,m)
  })(window,document,'script','https://www.google-analytics.com/analytics.js','ga');

  ga('create', 'UA-30013353-2', 'auto');
  ga('send', 'pageview');

</script>

