---
title: "2017 New Zealand Election"
output: 
  flexdashboard::flex_dashboard:
    theme: spacelab
    self_contained: false
    lib_dir: lib
    mathjax: null
    navbar:
      - { title: "Intro", icon: "", href: "", align: left }
      - { title: "Dumbbells", icon: "ion-ios-settings-strong", href: "dumbbells.html", align: left }
      - { title: "Hexmaps", icon: "fa-map", href: "hexmaps.html", align: left }
      - { title: "3D Hexmaps", icon: "fa-map", href: "3dhex.html", align: left }
      - { title: "@dakvid", icon: "fa-twitter", href: "https://twitter.com/dakvid", align: right }
    social: menu
    source_code: https://github.com/dakvid/election2017
---

```{r init, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```

```{r setup, include=FALSE}
library(flexdashboard)
```

## 

### Welcome

We just (September 2017) had an election in New Zealand, and try as I might I could not
resist trying to visualise some of the results. These are just what piqued my interest,
rather than an attempt to be complete or appeal to everyone. Hopefully others find them
interesting, if not useful.

David Friggens, October 2017

### Hexmaps

Geographical maps of results don't give a good representation of the spread of support
various parties have. So here are some cartograms with one tile per electorate. You can
see candidate race winners, vote concentration for all parties, and which electorates
favoured the incumbent coalition or not.

### Others

Unsurprisingly, other people have done some fantastic visualisations of results too.
If you like these, you'll like those too. (And if you don't like these, you may
like those better!)


## 

### Dumbbells

We know which parties gain and lost votes since 2014, but the national number doesn't
tell the full story. Was there universal swing, or were there localised ups and
downs? These dumbbell plots help make it clear.

### Desktop-friendly

I wanted to make these pages mobile friendly. However, whilst they don't "fail" as
such, they don't really succeed on a small screen. If you're like me you'll be
wanting to spend some time poking around with the mouse though, so it's probably
not a bad thing.

### 3D Hexmaps

I wanted to go further than the five colour shades in the hexmaps, so made these
hexmap / 3d barchart hybrids. They go give a bit more, but are potentially difficult
to interpret. I'm just happy that I managed to make it work!



<script>
  (function(i,s,o,g,r,a,m){i['GoogleAnalyticsObject']=r;i[r]=i[r]||function(){
  (i[r].q=i[r].q||[]).push(arguments)},i[r].l=1*new Date();a=s.createElement(o),
  m=s.getElementsByTagName(o)[0];a.async=1;a.src=g;m.parentNode.insertBefore(a,m)
  })(window,document,'script','https://www.google-analytics.com/analytics.js','ga');

  ga('create', 'UA-30013353-2', 'auto');
  ga('send', 'pageview');

</script>

