---
title: "2017 New Zealand Election (Provisional Results)"
output: 
  flexdashboard::flex_dashboard:
    theme: flatly
    self_contained: false
    lib_dir: lib
    mathjax: null
---

```{r init, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```

```{r setup, include=FALSE}
library(flexdashboard)
library(sf)
library(magrittr)
library(dplyr)
library(purrr)
library(stringr)
library(leaflet)
library(leaflet.minicharts)
library(nzelect)
library(htmltools)
library(htmlwidgets)

begin_label_text <- "<div style='font-size:12px;float:left'>"
begin_label_text_200 <- "<div style='font-size:12px;width:200px;float:left'>"
begin_label_header <- "<span style='font-size:18px;font-weight:bold'>"
end_label_header <- "</span><br>"
end_label_text <- "</div>"

load("data_geo/nzhex.rda")
load("data_votes/preliminary/GE2017_prelim.rda")

party_voters_2017 <-
  GE2017_prelim %>% 
  filter(VotingType == "Party",
         !str_detect(Party, "nformal")) %>% 
  group_by(id) %>% 
  summarise(Voters = sum(Votes)) %>% 
  ungroup()

candidate_voters_2017 <-
  GE2017_prelim %>% 
  filter(VotingType == "Candidate",
         !str_detect(Candidate, "nformal")) %>% 
  group_by(id) %>% 
  summarise(Voters = sum(Votes)) %>% 
  ungroup()

getLeafletOptions <- function(minZoom, maxZoom, ...) {
  leafletOptions(
    crs = leafletCRS("L.CRS.Simple"),
    minZoom = minZoom, maxZoom = maxZoom,
    dragging = TRUE, zoomControl = FALSE,
    tap = FALSE,
    attributionControl = FALSE , ...)
}

getFactorPal <- function(f) {
  colorFactor(colormap::colormap(
    colormap = colormap::colormaps$hsv,
    nshades = length(f)), f)
}
```

# Hexmaps {.storyboard}

### Hello HexMaps: a hexagonal cartogram of New Zealand electorates (2014-2017)

```{r hex_regions}
leaflet(options = getLeafletOptions(5.8, 5.8)) %>% 
  addPolygons(
    data = nzhex,
    label = ~nzhex %>% 
      transpose() %>% 
      map(~ paste0(begin_label_text,
               begin_label_header, .x$electorate_name, end_label_header,
               "In ", .x$region, ".",
               end_label_text) %>% HTML()),
    weight=2, color='#000000', group = 'Regions',
    fillOpacity = 0.6, opacity = 1, fillColor = ~getFactorPal(region)(region),
    highlightOptions = highlightOptions(weight = 4)
  ) %>% 
  addLabelOnlyMarkers(
    data = nzhex %>% st_centroid(),
    label = ~abbr,
    labelOptions = labelOptions(
      noHide = 'T', textOnly = T,
      offset=c(-6,-11))
  ) %>% 
  onRender(
    "function(el, t) {
      var myMap = this;
      // get rid of the ugly grey background
      myMap._container.style['background'] = '#ffffff';
      myMap.scrollWheelZoom.disable();
    }"
  )
```

***

This is a [cartogram](https://en.wikipedia.org/wiki/Cartogram) of the current
(2014-2017) New Zealand electorates, with one hexagon for each electorate.
Electorates have roughly equal populations, so this is a reasonable map of
population distribution. The North and South Islands should be obvious; the
M훮ori electorates make up a second mini NZ in the bottom right.

The colours indicate the regional groupings, to help orient you to the
distorted geography. Many electorates cross regional boundaries, so these
are just a rough guide. Most notably, the Wellington region is split between
several M훮ori electorates. (Credit for the regional groupings to
[Chris Knox in the Herald](http://insights.nzherald.co.nz/election/)).

The hexagonal layout is due to 
[Joseph Wright](http://mapdruid.blogspot.co.nz/2015/05/hexagonal-tile-map-of-new-zealand.html),
and I used his shapefiles as a base.

Chris McDowall made [a different hexagonal layout](http://hindsight.clerestories.com/2014/01/06/chris-mcdowall-hexagonal-maps/)
for the 2011 electorates, which [I adapted](http://david.frigge.nz/nzhex2014/) in 2014.
I think I like that better in terms of the geographic layout, but it is very
skinny so it's less compact for on-screen display.

Click through the storyboard and see different data from the election
visualised in hexmaps...


### Electorate Winners

```{r hex_candidates}
electorate_winners <-
  GE2017_prelim %>%
  filter(VotingType == "Candidate") %>%
  group_by(id, Party) %>%
  summarise(Votes = sum(Votes)) %>%
  top_n(1, Votes) %>%
  ungroup() %>% 
  inner_join(candidate_voters_2017) %>% 
  mutate(PerCent = round(Votes / Voters * 100, 1)) %>% 
  mutate(party_colour = case_when(
    Party == "National Party" ~ parties_v["National"],
    Party == "Labour Party" ~ parties_v["Labour"],
    Party == "New Zealand First Party" ~ parties_v["NZ First"],
    Party == "ACT New Zealand" ~ parties_v["ACT"],
    Party == "United Future" ~ parties_v["United Future"],
    Party == "M훮ori Party" ~ parties_v["M훮ori"]
  )) %>% 
  mutate(opa = 0.1 + 0.85 * ((PerCent - min(PerCent)) / (max(PerCent) - min(PerCent))))
electorate_winners$labeltext <-
  electorate_winners %>% 
  inner_join(nzhex) %>% 
  transpose() %>% 
  map(~ paste0(begin_label_text,
               begin_label_header, .x$electorate_name, end_label_header,
               .x$Party, " won with ", prettyNum(.x$Votes, big.mark=','), " votes.",
               "<br>",
               "<div style='width:95%'>",
               "<span style='color:", .x$party_colour, ";float:left'>", .x$PerCent, "%</span>",
               "<br clear='all' />",
               "<span style='background:", .x$party_colour, ";width:", .x$PerCent, "%;float:left'>&nbsp;</span>",
               "</div>",
               end_label_text) %>% 
        HTML())

leaflet(options = getLeafletOptions(5.8, 5.8)) %>% 
  addPolygons(
    data = inner_join(nzhex, electorate_winners),
    label = ~labeltext,
    weight=2, color='#000000', group = 'Electorate Seats',
    fillOpacity = ~opa, opacity = 1, fillColor = ~party_colour,
    highlightOptions = highlightOptions(weight = 4)
  ) %>% 
  addLabelOnlyMarkers(
    data = nzhex %>% st_centroid(),
    label = ~abbr,
    labelOptions = labelOptions(
      noHide = 'T', textOnly = T,
      offset=c(-6,-11))
  ) %>% 
  onRender(
    "function(el, t) {
      var myMap = this;
      // get rid of the ugly grey background
      myMap._container.style['background'] = '#ffffff';
    }"
  )
```

***

Each hexagon is coloured for the party of the winning candidate. The strength of colour tracks
the percentage of votes the winner received --- from
`r electorate_winners %>% top_n(-1, PerCent) %$% PerCent`% for
`r electorate_winners %>% top_n(-1, PerCent) %$% Party` in
`r electorate_winners %>% top_n(-1, PerCent) %>% inner_join(nzhex) %$% electorate_name`
to
`r electorate_winners %>% top_n(1, PerCent) %$% PerCent`% for
`r electorate_winners %>% top_n(1, PerCent) %$% Party` in
`r electorate_winners %>% top_n(1, PerCent) %>% inner_join(nzhex) %$% electorate_name`.


